Characterizing edges in signed and vector-valued graphs

### 2017-12-20 conclusion/main.tex
pass over conclusion

  \item In \autoref{chap:troll}~\autocite{trollSign17}, we first focused on directed signed graphs,
    as they model social interactions in several domains, such as e-commerce and collaborations.
    Here edges have two possible semantics, positive or negative. Our goal was to solve a supervised
    binary classification problem by predicting the sign of test edges. For that, we first
    introduced a simple sign generative model. It endows nodes with two parameters --trollness and
    unpleasantness--, governing their incoming and outgoing behavior. This might seem like a
    simplistic way of expressing the notion of partial homophily and heterophily. However, we
    presented two methods to approximate this Bayes predictor, suggesting a trade off between
    theoretical guarantees and input requirement. We showed experimentally that those methods are
    competitive with the state of the art, while being more scalable. Motivated by the regularity
    measure offered by trollness and unpleasantness, and the inability of existing methods to cope
    with new nodes joining the graph, we furthermore develop the first online algorithm for the
    \esp{} problem.

@@

  \item Finally, in \autoref{chap:vector}, we considered the case where there are more than two types
    of edge. This means that instead of signed graphs, our input shifted to multilayer graphs, which
    are also ubiquitous to describe real world phenomena. Moreover, we modified the learning
    setting, so that no edge label are available, but instead nodes carry a profile vector. However,
    our goal remained to predict edge type, in what could then be described as a clustering problem
    with side information and interpretability requirements. In order to do so, we first defined a
    similarity between nodes, which accommodate both heterophily and homophily. We then formulate
    the problem as simultaneously finding $k$ special vectors called directions, and assigning one
    of them to every edge in order to maximize a scalar \emph{goodness} function. After introducing
    baselines oblivious of the graph topology, we added constraints resulting in one vectorial and
    one matricial optimization objectives. Finally, we showed on synthetic data how those methods
    can recover the planted ground truth.

@@

\subsection{Reciprocal recommendation}

@@

Reciprocal recommendations present specific challenges~\autocite{Andrews2015}. Among them are
asymmetry and volume.  Asymmetry refers to the imbalance between the two groups, in terms of size
and bargaining power. For instance, depending of the economic situation and the field, there might
be a lot of job seekers for a few open positions, in which case employers can be very selective.
Likewise, it is generally observed that dating websites have a higher proportion of men, meaning
that women can afford to choose among numerous suitors~\autocite{TinderDesc16}. By volume, we mean
that users on both side can only handle a few recommendations at the same time. This is sharp
contrast with traditional user-item recommendation systems, where recommending simultaneously the
same item (say a movie) to thousands of users is not an issue. On the other hand, we have to keep
user engaged and prevent them from being idle. This is done by ensuring they are regularly the
subject or object of some recommendations.

As the name implies, reciprocal recommendation has often be seen as a special case of recommender
system~\autocite{Andrews2015}. Besides this paradigm, we could also cast it as an adaptive, active,
cost sensitive \esp{} problem on bipartite directed signed graphs with side information. Let us
decompose this last sentence, starting with the input graph. It is naturally bipartite, because of
the presence of two groups in which there is no internal recommendation. Furthermore, every time one
user expresses interest or disinterest in another user, this creates a directed signed edge. Such a
graph is thus a special case of the \dssn{} studied in \autoref{chap:troll}. Each nodes carries two
kind of information. The first is explicitly provided by the user, in order to arise interest in
users from the other group. The second is implicitly inferred from the user activity. In particular,
the interactions with other nodes allow us to build an implicit preference model for every user.
Combining these two sources of information yield a profile of each user, like in
\autoref{chap:vector}. Making recommendations in such a graph therefore boils down to finding non
existing positive edges. However, to cope with the two challenges presented in the previous
paragraph, we follow the setting of \autoref{chap:cc} and use an active approach. More precisely,
respecting the volume constraints would require an adaptive algorithm, which acquires the most
useful information without over- or under-loading any nodes. As for the group asymmetry, this could
be addressed by using a cost sensitive classification error, so that the error rate at a given node
is inversely proportional to its importance in the network.

As in any worthy machine learning endeavor, reliably evaluating and comparing approaches require
large amount of real data. This is particularly true in such an applied setting, but the most
realistic evaluation is to perform A/B testing on a live system with enough traffic to give some
statistical significance. This explains why most previous studies rely on the cooperation of
industrial partners~\autocites{Andrews2015}{Kille2015}{Krzywicki2015}{Xia2015}{Alanazi}. Thus, this
is also something we would like to do in the future.

@@

As an illustration in the context of this thesis, we assume that data was organized as a graph,
where nodes were possibly associated with a feature vector. However, there might an different
representation of those nodes that would better suit our needs. Indeed,
we mentioned already in \autoref{sec:troll_related} and \autoref{sec:edge_related} that a recent and
promising trend regarding learning in graph is node embedding. In addition to the references cited
earlier, this also witnessed by recent works on embedding attributed nodes, either in signed
graphs~\autocites{SignedFS17}{attributedSNE17}{SHINE18} or in general
graphs~\autocites{Liao2017}{AANE17}{DynamicEmbedding17}. At first sight, those methods seem
radically different from traditional graph methods. However, just like word embedding methods have
been linked to matrix
factorization~\autocites{WordEmbeddingAsFactorization14}{Hellinger14}{AroraRandWalk16}, so have been
popular and successful node embedding methods~\autocite{embeddingAsFactorization18}. These methods
are the simple ones that do not consider node attributes and multiple types of edges, such as
DeepWalk~\autocite{DeepWalk14}, LINE~\autocite{LINE15} and node2vec~\autocite{node2vec16}. For
instance, \textcite{embeddingAsFactorization18} show that the node embedding of DeepWalk on the
weighted graph $G$ are approximately the left singular vectors of the following matrix:
\begin{equation}
  \label{eq:deepwalk}
  \log \left( vol(G) \left( \frac{1}{T} \sum_{r=1}^T (D^{-1}A)^r\right) D^{-1} \right) - \log b
\end{equation}
where $vol(G)$ is the sum of all weighted degrees, $T$ is the context window size, $D$ is the
diagonal matrix of degrees, $A$ is the adjacency matrix of the graph and $r$ is the number of
negative samples in skip-gram. There are several ways such findings could be used in our context:
\begin{itemize}

>>>>

  \item In \autoref{chap:troll}~\autocite{trollSign17}, we first focused on directed signed graphs,
    as they model social interactions in several domains, such as e-commerce and collaborations.
    In such graphs, edges have two possible semantics, positive or negative. Our goal was to solve a supervised
    binary classification problem by predicting the sign of test edges. For that, we first
    introduced a simple sign generative model. It endows nodes with two parameters --trollness and
    unpleasantness--, governing their incoming and outgoing behavior. This might seem like a
    simplistic way of expressing the notion of partial homophily and heterophily. However, we
    presented two methods to approximate this Bayes predictor, suggesting a trade off between
    theoretical guarantees and input requirement. We showed experimentally that those methods are
    competitive with the state of the art, while being more scalable. Motivated by the regularity
    measure offered by trollness and unpleasantness, and the inability of existing methods to cope
    with new nodes joining the graph, we furthermore developed the first online algorithm for the
    \esp{} problem.

@@

  \item Finally, in \autoref{chap:vector}, we considered the case where there are more than two types
    of edge. This means that instead of signed graphs, our input shifted to multilayer graphs, which
    are also ubiquitous in describing real world phenomena. Moreover, we modified the learning
    setting, so that no edge label are available, but instead nodes carry a profile vector. However,
    our goal remained to predict edge type, in what could then be described as a clustering problem
    with side information and interpretability requirements. In order to do so, we first defined a
    similarity between nodes, which accommodates both heterophily and homophily. We then formulated
    the problem as simultaneously finding $k$ special vectors called directions, and assigning one
    of them to every edge in order to maximize a scalar \emph{goodness} function. After introducing
    baselines oblivious of the graph topology, we added constraints resulting in one vectorial and
    one matricial optimization objectives. Finally, we showed on synthetic data how those methods
    can recover the planted ground truth.

@@

\subsection{Reciprocal recommendation}
\label{ssec:conc_rr}

@@

Reciprocal recommendations present specific challenges~\autocite{Andrews2015}. Among them are
asymmetry and volume.  Asymmetry refers to the imbalance between the two groups, in terms of size
and bargaining power. For instance, depending of the economic situation and the field, there might
be a lot of job seekers for a few open positions, in which case employers can be very selective.
Likewise, it is generally observed that dating websites have a higher proportion of men, meaning
that women can afford to choose among numerous suitors~\autocite{TinderDesc16}. By volume, we mean
that users on both side can only handle a few recommendations at the same time. This is sharp
contrast with traditional user-item recommendation systems, where recommending simultaneously the
same item (say a movie) to thousands of users is not an issue. On the other hand, we have to keep
users engaged and prevent them from being idle. This is done by ensuring they are regularly the
subject or object of some recommendations.

As the name implies, reciprocal recommendation has often be seen as a special case of recommender
system~\autocite{Andrews2015}. Besides this paradigm, we could also cast it as an adaptive, active,
cost sensitive \esp{} problem on bipartite directed signed graphs with side information. Let us
decompose this last sentence, starting with the input graph. It is naturally bipartite, because of
the presence of two groups in which there is no internal recommendation. Furthermore, every time one
user expresses interest or disinterest in another user, this creates a directed signed edge. Such a
graph is thus a special case of the \dssn{} studied in \autoref{chap:troll}. Each nodes carries two
kind of (side) information. The first is explicitly provided by the user, in order to arise interest in
users from the other group. The second is implicitly inferred from the user activity. In particular,
the interactions with other nodes allow us to build an implicit preference model for every user.
Combining these two sources of information yield a profile of each user, like in
\autoref{chap:vector}. Making recommendations in such a graph therefore boils down to finding non
existing positive edges. However, to cope with the two challenges presented in the previous
paragraph, we follow the setting of \autoref{chap:cc} and use an active approach. More precisely,
respecting the volume constraints would require an adaptive algorithm, which acquires the most
useful information without over- or under-loading any nodes. As for the group asymmetry, this could
be addressed by using a cost sensitive classification error, so that the error rate at a given node
is inversely proportional to its importance in the network.

As in any worthy machine learning endeavor, reliably evaluating and comparing approaches require
large amount of real data. This is particularly true in such an applied setting, where the most
realistic evaluation is to perform A/B testing on a live system with enough traffic to deliver
statistical significance. This explains why most previous studies rely on the cooperation of
industrial partners~\autocites{Andrews2015}{Kille2015}{Krzywicki2015}{Xia2015}{Alanazi}. Thus, this
is also something we would like to do in the future.

@@

As an illustration in the context of this thesis, we assumed that data was organized as a graph,
where nodes were possibly associated with a feature vector. However, there might an different
representation of those nodes that would better suit our needs. Indeed,
we mentioned already in \autoref{sec:troll_related} and \autoref{sec:edge_related} that a recent and
promising trend regarding learning in graph is node embedding. In addition to the references cited
earlier, this also witnessed by recent works on embedding attributed nodes, either in signed
graphs~\autocites{SignedFS17}{attributedSNE17}{SHINE18} or in general
graphs~\autocites{Liao2017}{AANE17}{DynamicEmbedding17}. At first sight, those methods seem
radically different from traditional graph methods. However, just like word embedding methods have
been linked to matrix
factorization~\autocites{WordEmbeddingAsFactorization14}{Hellinger14}{AroraRandWalk16}, so have been
popular and successful node embedding methods~\autocite{embeddingAsFactorization18}. These methods
are the simple ones that do not consider node attributes and multiple types of edges, such as
DeepWalk~\autocite{DeepWalk14}, LINE~\autocite{LINE15} and node2vec~\autocite{node2vec16}. For
instance, \textcite{embeddingAsFactorization18} show that the node embedding of DeepWalk on the
weighted graph $G$ are approximately the left singular vectors of the following matrix:
\begin{equation}
  \label{eq:deepwalk}
  \log \left( vol(G) \left( \frac{1}{T} \sum_{r=1}^T (D^{-1}A)^r\right) D^{-1} \right) - \log b
\end{equation}
where $vol(G)$ is the sum of all weighted degrees, $T$ is the context window size, $D$ is the
diagonal matrix of degrees, $A$ is the adjacency matrix of the graph and $r$ is the number of
negative samples in skip-gram. There are several ways such findings could be used in our context:
\begin{itemize}
